A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-19 tex/body.tex
more edits

\begin{figure}[t!]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  t \in \tEc
}{
  \Ec := \Ec,\: t \teq t
}
\)
\rn{Refl}
\qquad
\(
\inferrule{
 t \teq u \in \Ec
}{
 \Ec := \Ec,\: u \teq t
}
\)
\rn{Sym}
\qquad
\(
\inferrule{
  s \teq t,\; t \teq u \in \Ec
}{
  \Ec := \Ec,\: s \teq u
}
\)
\rn{Trans}
\\[5\jot]
\(
\inferrule{
  \vec t \teq \vec u \in \Ec
  \betweenantes
  \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
}{
  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
}
\)
\rn{Cong}
\qquad
\(
\inferrule{
  t \teq u,\; t \tneq u \in \Ec
}{
  \bot
}
\)
\rn{Conflict}
\\[5\jot]
\(
\inferrule{
  \const{C}( \vec t \,) \teq \const{C}( \vec u ) \in \Ec
}{
  \Ec := \Ec,\: \vec t \teq \vec u
}
\)
\rn{Inject}
\qquad
\(
\inferrule{
  \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
  \betweenantes
  \const{C} \not= \const{D}
}{
  \bot
}
\)
\rn{Clash}
\end{tabular}
\vspace*{-3pt} %% TYPESETTING HACK
\caption{\,Derivation rules for bidirectional closure%.
}
\label{fig:cc-rules}
\end{figure}

@@

\begin{figure}[t!]
%\vspace*{+3pt}
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
%  t\typ{\delta}
%  \betweenantes
  \delta \in \Data
  \betweenantes
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \smash{\ec{t\typ{\delta}}} = \MU x.\; u
  \betweenantes
  x \in \FV( u )
}{
  \bot
}
\)
\rn{Acyclic}
\\[4\jot]
\(
\inferrule{
% t, u\typ{\delta}
% \betweenantes
 \delta \in \Codata
 \betweenantes
 %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
 \Val \smash{\ec{t\typ{\delta}}} \aequiv \Val \smash{\ec{u\typ{\delta}}}
}{
 \Ec := \Ec,\: t \teq u
}
\)
\rn{Unique}
\end{tabular}
\vspace*{-3pt} %% TYPESETTING HACK
\caption{\,Derivation rules for acyclicity and uniqueness%.
}
\label{fig:ab-rules}
\end{figure}

@@

\begin{theorem}[Refutation Soundness]%
\label{thm:rs}%
\afterDot
If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
\end{theorem}
\begin{proof}
The proof is by structural induction on the derivation tree with root node $\Ec$.
If the tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
then $\Ec$ is $\thD$-unsatisfiable.
For \rn{Conflict}, this is a consequence of equality reasoning.
For \rn{Clash}, this is a consequence of distinctness.
For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible.
If the child nodes of $\Ec$ are closed derivation trees
whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
$\thD$-unsatisfiable
for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
Since by exhaustiveness, all models of $\thD$ entail (exactly) one
$\DISC{\negvthinspace j}$,
$\Ec$ is $\thD$-unsatisfiable.
Otherwise, the child node of $\Ec$ is a closed derivation tree
whose root node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
In all these cases, $\Ec \models_{\thD} t \teq u$.
For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
For \rn{Inject}, this is a consequence of injectivity.
For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is one.
By the induction hypothesis, $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable
and thus $\Ec$ is $\thD$-unsatisfiable.
\qed
\end{proof}

@@

The $\mu$-term $u$
is a \emph{self-similar subterm} of~$t$ if
$u$ is a proper subterm of $t$,
$t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
$u$ is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
The term $t$ is \emph{normal} if it does not contain a self-similar subterm
and all of its proper subterms are also normal.
For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal
because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
Their arguments have the same expansion with respect to~$t$:
$\expand{t}{\MU y.\; \const{C}( y )} =
\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
\MU y.\; \const{C}( y )$
is $\alpha$-equivalent to
$\expand{t}{y} =
\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
\MU y.\; \const{C}( y )$.
The term $t = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $t$,
noting that
%
$\expand{t}{\MU y.\; \const{C}( x )}
 = \MU y.\; \const{C}\bigl( \expand{t}{x}^{\smash{\{y\}}} \bigr)
 = \MU y.\; \const{C}\bigl( \expand{t}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
 = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{t}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
 = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
is $\alpha$-equivalent to $\expand{t}{x} = t$.
% for the renaming $\{ x \mapsto y, y \mapsto x \}$.

For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$,
its \emph{normal form} $\nf{t}$ is obtained
by replacing all of the self-similar subterms of $t$ with $x$
%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
and by recursively normalizing the subterms.
For variables, $\nf{x} = x$.
Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.

@@

We now define the class of interpretations for $\thD$.
$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a set of
domain elements for that type.
$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
If $\const{f}$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
%To simplify the presentation,
All types are interpreted as sets of $\mu$-terms, but
only values of types in $\Codata$ may contain cycles.
%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
%a typical representation for these values can be obtained by
%dropping each of its $\mu$-bindings.
%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.

\begin{definition}[Normal Interpretation]
\afterDot%
\label{def:norm-interpretation}%
\rm
An interpretation $\J$ is \emph{normal} if these conditions are met:
\begin{enumerate}
\item
For each type $\tau$,
%%% sets *contain* elements, they *include* other sets
$\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
\item
For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
$\interp{\J}{\const{C}}( \interp{\J}{\vec t} )$ is the value
in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.
\item
For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
$\expand{\interp{\J}{t}}{u_k}$.
\end{enumerate}
\end{definition}

@@

\begin{theorem}[Solution Soundness]%
\label{thm:ss}%
\afterDot
If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
\end{theorem}
\begin{proof}
Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
%all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
%The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
We consider a normal interpretation $\J$
that interprets wrongly applied selectors based on equality information in $\Fc$
and that interprets the variables of $\Fc$ based on the completion $\ValC$ of the mapping $\Val$ from phase~2.
For the latter, let $\interp{\J}{x}$ be the value in $\J$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.

We first show $\J$ satisfies all equalities in $t_1 \teq t_2 \in \Fc$, where $t_1$ and $t_2$ have type $\tau$.
We show,
by structural induction on $t$,
that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
which since $\J$ is normal, implies $\J \models t_1 \teq t_2$.

@@

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \interp{\J}{u_1}, \ldots, \interp{\J}{u_n} ) }$ for some fresh $x$,
which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
Call this term $t_u$.
%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply, % to $\Fc$,
by the construction of $\ValC$ we have $\ValC \ec{t}$ is a term
of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
where
$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t_u$.
\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
$\expand{t_u}{u_\iPrime} \vsim \ValC \ec{u_i}$.
Thus $\expand{t_u}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
%By Lemma~\ref{lem:mu-cong}, we have $t_u \vsim t_w$.
Thus, $\interp{\J}{t} \vsim t_u \vsim \ValC \ec{t}$, and
we have $\interp{\J}{t} \vsim \ValC \ec{t}$.

>>>>

\begin{figure}[t!]
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
  t \in \tEc
}{
  \Ec := \Ec,\: t \teq t
}
\)
\rn{Refl}
\qquad
\(
\inferrule{
 t \teq u \in \Ec
}{
 \Ec := \Ec,\: u \teq t
}
\)
\rn{Sym}
\qquad
\(
\inferrule{
  s \teq t,\; t \teq u \in \Ec
}{
  \Ec := \Ec,\: s \teq u
}
\)
\rn{Trans}
\\[5\jot]
\(
\inferrule{
  \vec t \teq \vec u \in \Ec
  \betweenantes
  \const f( \vec t \,),\, \const f( \vec u ) \in \tEc
}{
  \Ec := \Ec,\: \const f( \vec t \,) \teq \const f( \vec u )
}
\)
\rn{Cong}
\qquad
\(
\inferrule{
  t \teq u,\; t \tneq u \in \Ec
}{
  \bot
}
\)
\rn{Conflict}
\\[5\jot]
\(
\inferrule{
  \const{C}( \vec t \,) \teq \const{C}( \vec u ) \in \Ec
}{
  \Ec := \Ec,\: \vec t \teq \vec u
}
\)
\rn{Inject}
\qquad
\(
\inferrule{
  \const{C}( \vec t \,) \teq \const{D}( \vec u ) \in \Ec
  \betweenantes
  \const{C} \not= \const{D}
}{
  \bot
}
\)
\rn{Clash}
\end{tabular}
%\vspace*{-3pt} %% TYPESETTING HACK
\caption{\,Derivation rules for bidirectional closure%.
}
\label{fig:cc-rules}
\end{figure}

@@

\begin{figure}[t!]
%\vspace*{+3pt}
\normalsize
\centering
\begin{tabular}{c}
\(
\inferrule{
%  t\typ{\delta}
%  \betweenantes
  \delta \in \Data
  \betweenantes
  %\ttpath{\Ec}{\ec{t}}{\ec{t}} \neq \emptyset
  \Val \smash{\ec{t\typ{\delta}}} = \MU x.\; u
  \betweenantes
  x \in \FV( u )
}{
  \bot
}
\)
\rn{Acyclic}
\\[4\jot]
\(
\inferrule{
% t, u\typ{\delta}
% \betweenantes
 \delta \in \Codata
 \betweenantes
 %\tpath{\Ec}{\ec{t}} = \tpath{\Ec}{\ec{u}}\sigma \neq \emptyset
 \Val \smash{\ec{t\typ{\delta}}} \aequiv \Val \smash{\ec{u\typ{\delta}}}
}{
 \Ec := \Ec,\: t \teq u
}
\)
\rn{Unique}
\end{tabular}
%\vspace*{-3pt} %% TYPESETTING HACK
\caption{\,Derivation rules for acyclicity and uniqueness%.
}
\label{fig:ab-rules}
\end{figure}

@@

\begin{theorem}[Refutation Soundness]%
\label{thm:rs}%
\afterDot
If there exists a closed derivation tree with root node\/ $\Ec$, then\/ $\Ec$ is $\thD$-unsatisfiable.
\end{theorem}
\begin{proof}
The proof is by structural induction on the derivation tree with root node $\Ec$.
If the tree is an application of \rn{Conflict}, \rn{Clash}, or \rn{Acyclic},
then $\Ec$ is $\thD$-unsatisfiable.
For \rn{Conflict}, this is a consequence of equality reasoning.
For \rn{Clash}, this is a consequence of distinctness.
For \rn{Acyclic}, the construction of $\Val$ indicates that the class of values that term~$t$ can take in models of $\Ec$ is infeasible.
If the child nodes of $\Ec$ are closed derivation trees
whose root nodes are the result of applying \rn{Split} on $t\typ{\delta}$,
by the induction hypothesis $\Ec \mathrel\cup t \teq \smash{\const C_{\negvthinspace j}\bigl( \const s^1_{\negvthinspace j}( t ), \ldots, \const s^{\,\smash{n_{\negvthinspace j}}}_{\negvthinspace j}( t ) \bigr)}$ is
$\thD$-unsatisfiable
for each $\const C_{\negvthinspace j} \in \smash{\Ctr^\delta}$.
Since by exhaustiveness, all models of $\thD$ entail exactly one
$\DISC{\negvthinspace j}$,
$\Ec$ is $\thD$-unsatisfiable.
Otherwise, the child node of $\Ec$ is a closed derivation tree
whose root
node $\Ec \mathrel\cup t \teq u$ is obtained by applying one of the rules \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, \rn{Inject}, \rn{Unique}, or \rn{Single}.
In all these cases, $\Ec \models_{\thD} t \teq u$.
For \rn{Refl}, \rn{Sym}, \rn{Trans}, \rn{Cong}, this is a consequence of equality reasoning.
For \rn{Inject}, this is a consequence of injectivity.
For \rn{Unique}, the construction of $\Val$ indicates that the values of $t$ and $u$ are equivalent in all models of $\Ec$.
For \rn{Single}, $t$~and~$u$ must have the same value since the cardinality of their type is one.
By the induction hypothesis, $\Ec \mathrel\cup t \teq u$ is $\thD$-unsatisfiable
and thus $\Ec$ is $\thD$-unsatisfiable.
\qed
\end{proof}

@@

The $\mu$-term $u$
is a \emph{self-similar subterm} of~$t$ if
$u$ is a proper subterm of $t$,
$t$ is of the form $\MU x.\; \const{C}( t_1, \ldots, t_n )$,
$u$ is of the form $\MU y.\; \const{C}( u_1, \ldots, u_n )$,
and $\expand{t}{ t_k } \aequiv \expand{t}{ u_k }$ for all $k$. % \in [n]
The term $t$ is \emph{normal} if it does not contain a self-similar subterm
and all of its proper subterms are also normal.
For example, $t = \MU x.\; \const{C}( \MU y.\; \const{C}( y ) )$ is abnormal
because $\MU y.\; \const{C}( y )$ is a self-similar subterm of $t$.
Their arguments have the same expansion with respect to~$t$:
$\expand{t}{\MU y.\; \const{C}( y )} =
\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
\MU y.\; \const{C}( y )$
is $\alpha$-equivalent to
$\expand{t}{y} =
\MU y.\; \const{C}\bigl( \expand{t}{y}^{\smash{\{y\}}} \bigr) =
\MU y.\; \const{C}( y )$.
The term $u = \MU x.\; \const{C}( \MU y.\; \const{C}( x ) )$ is also abnormal,
since $\MU y.\; \const{C}( x )$ is a self-similar subterm of $u$,
noting that
%
$\expand{u}{\MU y.\; \const{C}( x )}
 = \MU y.\; \const{C}\bigl( \expand{u}{x}^{\smash{\{y\}}} \bigr)
 = \MU y.\; \const{C}\bigl( \expand{u}{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) ) }^{\smash{\{y\}}} \bigr)
 = \MU y.\; \const{C}\bigl( \MU x.\; \const{C}( \expand{u}{\MU y.\; \const{C}( x )}^{\smash{\{x,y\}}} ) \bigr)
 = \MU y.\allowbreak\; \const{C}( \MU x.\allowbreak\; \const{C}( y ) )$
is $\alpha$-equivalent to $\expand{u}{x} = u$.
% for the renaming $\{ x \mapsto y, y \mapsto x \}$.

For any $\mu$-term $t$ of the form $\MU x.\; \const C( \vec u )$,
its \emph{normal form} $\nf{t}$ is obtained
by replacing all of the self-similar subterms of $t$ with $x$
%and by replacing all of the abnormal subterms of $t$ with their corresponding normal forms.
and by recursively normalizing the other subterms.
For variables, $\nf{x} = x$.
Thus, $\nf{\MU x.\; \const{C}( \MU y.\; \const{C}( x ) )} = \MU x.\; \const{C}( x )$.

@@

We now define the class of interpretations for $\thD$.
$\interp{\J}{\tau}$ denotes the interpretation type $\tau$ in $\J$---that is, a
nonempty set of domain elements for that type.
$\interp{\J}{\const{f}}$ denotes the interpretation of a function $\const{f}$ in $\J$.
If $\const{f}$ is of type $ \tau_1 \times \cdots \times \tau_n \to \tau$,
then $\interp{\J}{\const{f}}$ is a total function from $\interp{\J}{\tau_1} \times \cdots \times \interp{\J}{\tau_n}$ to $\interp{\J}{\tau}$.
%To simplify the presentation,
All types are interpreted as sets of $\mu$-terms, but
only values of types in $\Codata$ may contain cycles.
%Since all values of types in $\Data$ and $\Nondata$ are acyclic,
%a typical representation for these values can be obtained by
%dropping each of its $\mu$-bindings.
%For example, the $\mu$-term $\MU x.\; \const{Z}$ representing the integer constant zero
%is simply $\const{Z}$ after dropping the $\mu$-binding of $x$.

\begin{definition}[Normal Interpretation]
\afterDot%
\label{def:norm-interpretation}%
\rm
An interpretation $\J$ is \emph{normal} if these conditions are met:
\begin{enumerate}
\item
For each type $\tau$,
%%% sets *contain* elements, they *include* other sets
$\interp{\J}{\tau}$ includes a maximal set of closed normal $\mu$-terms of that type that are
unique up to $\alpha$-equivalence and acyclic if $\tau \not\in \Codata$.
\pagebreak[2] %% TYPESETTING
\item
For each constructor term $\const{C}( \vec t \,)$ of type $\tau$,
$\interp{\J}{\const{C}}( \interp{\J}{\vec t} )$ is the value
in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
$\nf{\MU x.\; \const{C}( \interp{\J}{\vec t} )}$, where $x$ is fresh.
\item
For each selector term $\const s^k_{\negvthinspace j}( t )$ of type $\tau$, % for some $\vec u$,
if $\interp{\J}{t}$ is $\MU x.\; \const{C}_{\negvthinspace j}( \vec u )$, 
then $\J\negvvthinspace( \const s^k_{\negvthinspace j} )\bigl( \interp{\J}{t} \bigr)$ is the value
in $\interp{\J}{\tau}$ that is $\alpha$-equivalent to
$\expand{\interp{\J}{t}}{u_k}$.
\end{enumerate}
\end{definition}

@@

\begin{theorem}[Solution Soundness]%
\label{thm:ss}%
\afterDot
If there exists a derivation tree with root node $\Ec$ containing a saturated node, then $\Ec$ is $\thD$-satisfiable.
\end{theorem}
\begin{proof}
Let $\Fc$ be a saturated node in a derivation tree with root node $\Ec$.
%We will construct a model $\J$ for a set of equalities $\Fc$ that is equivalent to $\Fc_0$, and where
%all equivalence classes of $\Fc$ contain at least one variable, and all selectors in $\Fc$ are applied to variables only.
%The former comes with no loss of generality since new equalities of the form $y \teq t$ for fresh variable $y$ can be added to $\Fc$ without affecting its satisfiability,
%The latter also comes with no loss of generality since nested applications of selectors in $\Fc_0$ can be replaced by fresh variables while adding equalities to $\Fc$.
We consider a normal interpretation $\J$
that interprets wrongly applied selectors based on equality information in $\Fc$
and that interprets the variables of $\Fc$ based on the completion $\ValC$.
% of %the mapping $\Val$. % from phase~2.
For the variables, let $\interp{\J}{x\typ{\tau}}$ be the value in $\J(\tau)$ that is $\alpha$-equivalent with $\ValC \ec{x}$ for each variable $x \in \Terms(\Fc)$,
which by Lemma~\ref{lem:interpretation-completion}\PART{1} is guaranteed to exist.

We first show that $\J$ satisfies all equalities $t_1 \teq t_2 \in \Fc$. % where $t_1$ and $t_2$ have type $\tau$.
To achieve this, we show
by structural induction on $t\typ\tau$
that $\interp{\J}{t} \aequiv \ValC \ec{t}$ for all terms $t \in \Terms( \Fc )$,
which implies $\J \models t_1 \teq t_2$ since $\J$ is normal.

@@

If $t$ is a constructor term of the form $\const{C}( u_1, \ldots, u_n )$,
then $\interp{\J}{t}$ is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \interp{\J}{u_1}, \ldots, \interp{\J}{u_n} ) }$ for some fresh $x$,
which by the induction hypothesis is $\alpha$-equivalent with $\nf{\MU x.\; \const{C}( \ValC \ec{ u_1 }, \ldots, \ValC \ec{ u_n } )}$.
Call this term $t'$.
%By Lemma~\ref{lem:interpretation-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{t}$ is a term
%of the form $\MU {\Varec{t}}.\; \const{C}( w_1, \ldots, w_n )$
%where $\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
by the construction of $\ValC$ we have that $\ValC \ec{t}$ is a term
of the form $\MU {\Varec{t}}.\; \const{C}( w_1,\allowbreak \ldots,\allowbreak w_n )$
where
$\muvar(w_i) = {\Varec{u_i}}$ for each $i$.  Thus by Lemma~\ref{lem:interpretation-completion}\PART{2},
$\expand{\ValC \ec{t}}{w_i} = \ValC \ec{u_i}$.
For each $i$, let $u_\iPrime$ be the $i$\vvthinspace th argument of $t'$.
\begin{rep}By Lemma~\ref{lem:mu-norm-arg},\end{rep}%
\begin{conf}Clearly,\end{conf} %% TODO: Fix! ;)
$\expand{t'}{u_\iPrime} \vsim \ValC \ec{u_i}$.
Thus $\expand{t'}{u_\iPrime} \vsim \expand{\ValC \ec{t}}{ w_i }$. % for each $i \in [n]$.
%By Lemma~\ref{lem:mu-cong}, we have $t' \vsim t_w$.
Thus, $\interp{\J}{t} \vsim t' \vsim \ValC \ec{t}$, and
we have $\interp{\J}{t} \vsim \ValC \ec{t}$.
